---
title: "STAT/MATH 495: Problem Set 03"
author: "Kiryu Kawahata"
date: "2017-09-26"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
    collapsed: false
    smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=8, fig.height=4.5)

# Load packages
library(tidyverse)
library(broom)
library(mosaic)
library(splines)
library(gridExtra)
data1 <- read_csv("data/data1.csv")
data2 <- read_csv("data/data2.csv")
```


# Question

For both `data1` and `data2` tibbles (a tibble is a data frame with some
[metadata](https://blog.rstudio.com/2016/03/24/tibble-1-0-0#tibbles-vs-data-frames) attached):

* Find the splines model with the best out-of-sample predictive ability.
* Create a visualizaztion arguing why you chose this particular model.
* Create a visualizaztion of this model plotted over the given $(x_i, y_i)$ points for $i=1,\ldots,n=3000$.
* Give your estimate $\widehat{\sigma}$ of $\sigma$ where the noise component $\epsilon_i$ is distributed with mean 0 and standard deviation $\sigma$.

#Preview
A quick preview of what is in each data set:
```{r, echo=TRUE, warning=FALSE, message=FALSE}
favstats(data1$y)
favstats(data2$y)


ggplot(data = data1, aes(x = y)) + geom_density(adjust=0.4) + labs(title="data1 Exploration", x="y of the data") + theme(legend.position="none")

ggplot(data = data2, aes(x = y)) + geom_density(adjust=0.4) + labs(title="data2 Exploration", x="y of the data") + theme(legend.position="none")
```

# Data 1

```{r, echo=TRUE, warning=FALSE, message=FALSE}
Shuff_1 <- sample(seq_len(nrow(data1)), size = 1500)

Train <- data1[Shuff_1,]
Valid <- data1[-Shuff_1,]

#Below is a fuction that can easily produce spline models using different degrees of freedom. 

Model_by_df <- function(df) {
  Spline_Model_data1 <- smooth.spline(x = Train$x, y = Train$y, df=df)
  return(Spline_Model_data1)
}

Model_by_df_tidy <- function(df) {

Spline_Model_data1 <- smooth.spline(x = Train$x, y = Train$y, df=df)
Spline_data1_tidy <- Spline_Model_data1 %>% 
  broom::augment()
return(Spline_data1_tidy)
}


#For data1 I have chosen a model with 15 knots based on the plot as there is little under or overfitting with this model. 


data1_plot <- ggplot(Model_by_df_tidy(50), aes(x=x)) + 
  geom_point(aes(y=y)) + geom_line(aes(y=.fitted), col= "Red") + labs(title = "50 D of F data1 Spline Model Plot")
data1_plot_2 <- ggplot(Model_by_df_tidy(30), aes(x=x)) + 
  geom_point(aes(y=y)) + geom_line(aes(y= .fitted), col= "Blue") + labs(title = "30 D of F data1 Spline Model Plot") 
data1_plot_3 <- ggplot(Model_by_df_tidy(15), aes(x=x)) + 
  geom_point(aes(y=y)) + geom_line(aes(y= .fitted), col= "Purple") + labs(title = "15 D of F data1 Spline Model Plot")
data1_plot_4 <- ggplot(Model_by_df_tidy(10), aes(x=x)) + 
  geom_point(aes(y=y)) + geom_line(aes(y= .fitted), col= "Green") + labs(title = "10 D of F data1 Spline Model Plot")

grid.arrange(data1_plot, data1_plot_2, data1_plot_3, data1_plot_4, nrow = 2, ncol=2)


#Calculation of the the Root Mean Square Error(RMSE):

Trainset_test_data1 <- predict(Model_by_df(15), Train$x)

data1_t_RMSE <- sqrt(sum(Trainset_test_data1$y^2))/3000

validation_test_data1 <- predict(Model_by_df(15), Valid$x)


data1_V_RMSE <- sqrt(sum(validation_test_data1$y^2))/3000

data1_score_hat <- (data1_t_RMSE + data1_V_RMSE)/2

data1_score_hat 


```

Score-hat from the average RMSE for data1 is 0.0631 using k = 2 fold crossvalidation.


# Data 2

```{r, echo=TRUE, warning=FALSE, message=FALSE}

#Sampling/splitting of data 2 into train and validation sets. 

Shuff_2 <- sample(seq_len(nrow(data2)), size = 1500)

Train_2 <- data2[Shuff_2,]
Valid_2 <- data2[-Shuff_2,]

#Again easy functions to call models based on df/knots. Modified for new train sets to avoid reissue of unrelated data1 numbers. 

Model_by_df_2 <- function(df) {
  Spline_Model_data2 <- smooth.spline(x = Train_2$x, y = Train_2$y, df=df)
  return(Spline_Model_data2)
}

Model_by_df_tidy_2 <- function(df) {

Spline_Model_data2 <- smooth.spline(x = Train_2$x, y = Train_2$y, df=df)
Spline_data2_tidy <- Spline_Model_data2 %>% 
  broom::augment()
return(Spline_data2_tidy)
}


#This time for data2 I have decided to test a model with 17 knots in the spline model based on plot observations. 

data2_plot <- ggplot(Model_by_df_tidy_2(50), aes(x=x)) + 
  geom_point(aes(y=y)) + geom_line(aes(y= .fitted), col= "blue") + labs(title = "50 D of F data2 Spline Model Plot")
data2_plot_2 <- ggplot(Model_by_df_tidy_2(30), aes(x=x)) + 
  geom_point(aes(y=y)) + geom_line(aes(y= .fitted), col= "red") + labs(title = "30 D of F data2 Spline Model Plot")
data2_plot_3 <- ggplot(Model_by_df_tidy_2(15), aes(x=x)) + 
  geom_point(aes(y=y)) + geom_line(aes(y= .fitted), col= "green") + labs(title = "15 D of F data2 Spline Model Plot")
data2_plot_4 <- ggplot(Model_by_df_tidy_2(17), aes(x=x)) + 
  geom_point(aes(y=y)) + geom_line(aes(y= .fitted), col= "purple") + labs(title = "17 D of F data2 Spline Model Plot")

grid.arrange(data2_plot, data2_plot_2, data2_plot_3, data2_plot_4, nrow = 2, ncol=2)

#Calculation of the the Root Mean Square Error(RMSE):

Trainset_test_data2 <- predict(Model_by_df_2(15), Train_2$x)

data2_t_RMSE <- sqrt(sum(Trainset_test_data2$y^2))/3000

validation_test_data2 <- predict(Model_by_df_2(15), Valid_2$x)


data2_V_RMSE <- sqrt(sum(validation_test_data2$y^2))/3000

data2_score_hat <- (data2_t_RMSE + data2_V_RMSE)/2

data2_score_hat 



```

The RMSE score for data2 is 0.076 using k = 2 fold crossvalidation. I do however want to create a crossvalidation with at least 5 folds. As of writing I am trying to figure this out, and will include this if I can complete before pull. 
